\documentclass [11pt, a4paper]{report}
\pagestyle{plain}
\usepackage{nomencl}
\usepackage{graphicx}
\usepackage{url}
\usepackage[pdftex]{hyperref}
\usepackage{cite}
\usepackage[applemac]{inputenc}
\usepackage[
acronym,      %ein Abkürzungsverzeichnis erstellen
toc,          %Einträge im Inhaltsverzeichnis
section]      %im Inhaltsverzeichnis auf section-Ebene erscheinen
{glossaries}

\makeglossaries

\newacronym{rf}{RF}{Radiofrequency}
\newacronym{rfa}{RFA}{Radiofrequency Ablation}
\newacronym{ca}{CA}{Cryoablation}
\newacronym{hcc}{HCC}{Hepatocellular Carcinoma}
\newacronym{dicom}{DICOM}{Digital Imaging and Communication in Medicine}
\newacronym{mrml}{MRML}{Medical Reality Markup Language}
\newacronym{us}{US}{Ultrasound}
\newacronym{mri}{MRI}{Magnetic Resonance Imaging}
\newacronym{ct}{CT}{Computed Tomography}
\newacronym{mw}{MW}{Microwave}
\newacronym{mwa}{MWA}{Microwave Ablation}
\newacronym{fea}{FEA}{Finite Element Analysis}
\newacronym{litt}{LITT}{Laster-induced Thermotherapy}
\newacronym{hifu}{HIFU}{High-intensity focused ultrasound}
\newacronym{itk}{ITK}{The Insight Segmentation and Registration Toolkit}
\newacronym{vtk}{VTK}{The Visualization Toolkit}
\newacronym{gui}{GUI}{Graphical User Interface}
\newacronym{api}{API}{Application Programming Interface}
\newacronym{xml}{XML}{Extensible Markup Language}


\begin{document}

\include{frontmatter}
% *************** Table of contents ***************
\pagenumbering{roman}
\pagestyle{plain}
\tableofcontents

% *************** End of front matter ***************

\pagestyle{plain}

\pagenumbering{arabic}

\chapter*{Abstract}

\section*{Background and Goal}
Thermal ablation as a complementary treatment for localized tumors has been proving itself for the last years. Using thermal ablation, different techniques (e.g. Radiofrequency ablation, Cryoablation etc.) are applied to destroy cancer tissue either by heating or freezing. These techniques are applied in a minimal-invasive manner in contrast to surgical resection. Medical imaging modalities, such as Magnetic resonance imaging, computed tomography and ultrasound are essential due to minimal-invasiveness.
\\
Although thermal ablation is particularly used for the destruction of small tumors worldwide, metastases still re-appear more frequently than in case of surgical resection. Successful therapy depends on many factors. Despite of the additional removal of a 1cm sized margin around the tumor, metastases still re-occur sometimes. Incomplete removal of the tumor is mostly the reason for this circumstance.
\\ 
The collaboration of different experts and techniques suggests, that the operators would benefit from a software-platform for supporting the operators during the planning-phase. The software enables the abstract simulation of the intervention on basis of the patient's specific anatomy. Multiple applicators can be placed into the virtual body and the visualization of the resulting lesion bases upon data of the respective device vendors. The operators are enabled to see which structures lie in the trajectory of the applicator and above all which anatomical structures are adjacent to the tumor. Large vessels adjacent to the tumor will most likely limit the dimension of the ablation zone because of the blood flow. Those and other effects will be detected and avoided by applying the planning-software.

\section*{Methods and Results}

Before developing the prototype, the processes of the thermal ablation intervention must be understood. Different ablation techniques are described in this thesis. Devices of variable vendors were evaluated to determine the different ablation zone shapes. The requirements of such a planning-system are based upon the research work of Dr. Wolfgang Schramm.
\\
The software was developed as a module-extension of the Slicer software framework. The modular structure of Slicer allows the extension of its own functionality by adding different module types.  In the process, a Python scripted module was developed to implement the functionality of such a planning-software as a prototype. The resulting software is open source and free to access at https://github.com/HaichaoMiao/Slicer4-Module.

\chapter*{Zusammenfassung}

\section*{Hintergrund und Ziele}

Thermische Ablation als eine komplementäre Behandlung für lokale Tumore hat sich seit Jahren bewährt und wird immer häufiger eingesetzt. Dabei werden unterschiedliche Techniken (wie Hochfrequenzablation, Kryoablation etc.) angewendet, um das Krebsgewebe entweder durch Erhitzung oder Erfrierung zu zerstören. Die Behandlung erfolgt minimal-invasiv im Gegensatz zum chirurgischen Eingriff, wobei bildgebende Verfahren wie Magnetresonanztomographie, Computertomographie und Ultraschall unverzichtbar sind. 
\\
Obwohl thermische Ablation vor allem für die Behandlung von kleineren Tumoren immer häufiger eingesetzt wird, ist das Wiederauftreten von Metastasen jedoch häufiger als bei der chirurgischen Resektion. Der Erfolg von thermischer Ablation ist von vielen Faktoren abhängig. Trotz der zusätzlichen Entfernung eines 1cm Abstandes um den Tumor, kommt es häufig zu Wiederauftreten von Metastasen. Die Ursache dafür ist oft eine unvollständige Entfernung.
\\
Die Multidisziplinarität dieser Behandlung erfordert den Einsatz mehrerer Experten bzw. Techniken. Es ist daher naheliegend eine Software-Plattform zu entwickeln, welche die Operatoren vor allem während der Planungs-Phase unterstützt. Die Software ermöglicht eine abstrakte Simulation des Vorganges aufgrund der patientenspezifischen Anatomie. Mehrere Applikatoren können in den virtuellen Körper eingebracht und die resultierenden Ablationszonen aufgrund der Angaben des Gerätherstellers visualisiert werden. Die Operatoren haben dadurch die Möglichkeit zu erkennen, welche Organe und Knochen in der Einstichsbahn des Applikators liegen und vor allem, welche anatomischen Strukturen am Tumor anliegen. Insbesondere große Gefäße nahe am Tumor können das Ausmaß der Ablationszone beschränken, da diese die thermische Energie ableiten. Diese und andere Effekte können mit der Planungssoftware im voraus erkannt und vermieden werden. 

\section*{Methoden und Ergebnisse}

Für die Entwicklung des Prototyps musste zuerst der Vorgang der thermischen Ablation verstanden werden. Unterschiedliche Ablationstechniken wurden abgebildet und beschrieben. Geräte verschiedener Hersteller wurden ausgewertet, um die unterschiedlichen Formen der Ablationszone zu bestimmen. Die Anforderung an ein solches Planungssystems basiert insbesondere auf der Dissertation von Dr. Wolfgang Schramm. 
\\
Die Software wurde als Modul-Erweiterung des Slicer Software-Frameworks realisiert. Der modulare Aufbau von Slicer erlaubt die Erweiterung seiner eigenen Funktionalität mit unterschiedlichen Modultypen. Dabei wurde ein "Python scripted" Modul entwickelt, um die Funktionalität einer solchen Planungs-Software als Prototyp zu implementieren. Die fertig entwickelte Software ist open source und unter https://github.com/HaichaoMiao/Slicer4-Module frei zugänglich. 

\include{introduction}

\include{chapter1}

\include{chapter2}

\include{chapter3}

%\chapter{Outlook and Discussion}
%TODO
\chapter{Conclusio}
Over the past years, different thermal ablation techniques have been applied successfully to treat tumors. The treatment of malignant liver tumor, kidney tumor and many other localized tumor forms by using thermal energy sources has gained clinical acceptance. Although 
tumor patients benefit from the thermal ablation interventions, the technique itself is still limited by the maximum size of ablation zone that can be achieved. Thermal ablation is not an alternative to surgical resection but a complementary treatment form. 
\\
Before the intervention starts, many decisions have to be made, which have great impact on the outcome. Because of the multidiscipline nature of the treatment form, a planning phase may have major influence on the intervention. Many errors during the procedure, such as wrong applicator placement, could afflict irreversible damage to the patient. The software supports the operators during the planning phase by visualizing the procedure and making an adequate forecast of the treatment outcome. This helps the collaborators to picture and understand the intervention as a whole and thus the prototype helps avoiding errors. 
\\
The prototype gives visual feedback whether the trajectories of the placed probes are applicable or not. Additionally the prototype provides information of the coverage of tumor by the ablation zone on based upon the respective vendor or literature. Although the developed prototype is far from the actual application in the clinical practice as discussed before, it makes another step towards the goal of supporting the thermal ablation intervention by software.

\chapter{Terms and abbreviations}

%Abkürzungen ausgeben
\glsaddall
\printglossary[type=\acronymtype,style=list]

% \appendix
% include appendexes here

\include{appendix}

\listoffigures

\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}


